### 2025-03-22 latex/paper.tex
tex init

\documentclass{article}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{geometry}
\geometry{legalpaper, portrait, margin=.5in}
\title{CS 2051 Group Project}
\author{}
\date{March 2025}

\begin{document}

\maketitle
\end{document}

### 2025-03-23 c/main.cpp
Added Neural Network structure and forward pass implemenation

@@ -1,3 +1,4 @@
+#include "src/network/network.hpp"
 #include <iostream>
 
 int main() { std::cout << "Hello" << std::endl; }

### 2025-03-23 c/src/network/Optimizer.hpp

@@ -1,0 +1,14 @@
+#pragma once
+#include "Eigen/Dense"
+#include "types.hpp"
+
+// Base class for optimizers
+class Optimizer {
+public:
+    // Updates the given weights and biases from their gradients.
+    virtual void update(EigenMat &weights, EigenVec &biases,
+                        const EigenMat &weight_gradients,
+                        const EigenVec &bias_gradients,
+                        const EigenVec &activations) = 0;
+    virtual ~Optimizer() = default;
+};

### 2025-03-23 c/src/network/network.cpp

@@ -1,0 +1,123 @@
+#pragma once
+#include "network.hpp"
+#include "Eigen/Dense"
+#include "types.hpp"
+#include <vector>
+#include <stdint.h>
+#include <random>
+
+Network::Network(uint32_t layers, std::vector<uint32_t> layer_sizes,
+                 std::vector<float (*)(float)> activation_funcs,
+                 std::vector<float (*)(float)> activation_derivatives,
+                 OutputActivationFunc output_activation, CostFunc cost,
+                 OutputLayerDerivative output_layer_derivative)
+{
+    assert(
+        layers >= 2 &&
+        "network must have at least 2 layers"
+    );
+    assert(
+        layer_sizes.size() == layers &&
+        "mismatch between layers and layer_sizes"
+    );
+    for (int size : layer_sizes) {
+        assert(size > 0 && "invalid layer size of 0");
+    }
+    assert(
+        activation_funcs.size() == layers - 1 &&
+        "mismatch between layers and activations"
+    );
+    for (float (*activation)(float) : activation_funcs) {
+        assert(activation != NULL && "invalid null activation");
+    }
+    assert(output_activation != NULL && "invalid null output activation");
+    assert(cost != NULL && "invalid null cost function");
+    assert(output_layer_derivative != NULL && "invalid null output derivative");
+    this->layers = layers;
+    this->layer_sizes = layer_sizes;
+    this-> activation_funcs = activation_funcs;
+    this->activation_derivatives = activation_derivatives;
+    this->output_activation = output_activation;
+    this->cost = cost;
+    this->output_layer_derivative = output_layer_derivative;
+    this->activations.reserve(layers - 1);
+    this->weights.reserve(layers - 1);
+    this->biases.reserve(layers - 1);
+    // Randomly init weights and biases
+    std::random_device rd;
+    std::mt19937 gen(rd());
+    std::uniform_real_distribution<> dis(-1.0, 1.0);
+    for (int i = 0; i < layers - 1; i++) {
+        EigenMat current_weights(layer_sizes[i+1], layer_sizes[i]);
+        EigenMat current_biases(layer_sizes[i+1], layer_sizes[i]);
+        for (int col = 0; col < current_weights.cols(); col++) {
+            for (int row = 0; row < current_weights.rows(); row++) {
+                current_weights(row, col) = dis(gen);
+            }
+        }
+        for (int row = 0; row < current_biases.size(); row++) {
[…]

### 2025-03-23 c/src/network/network.hpp

@@ -1,0 +1,66 @@
+#pragma once
+#include "Eigen/Dense"
+#include "types.hpp"
+#include <stdint.h>
+#include <vector>
+
+// Forward declaration of Optimizer class
+class Optimizer;
+
+class Network {
+  public:
+    // Initializes a new Network from the given parameters.
+    Network(uint32_t layers, std::vector<uint32_t> layer_sizes,
+            std::vector<float (*)(float)> activation_funcs,
+            std::vector<float (*)(float)> activation_derivatives,
+            OutputActivationFunc output_activation, CostFunc cost,
+            OutputLayerDerivative output_layer_derivative);
+
+    // Feeds the given input through the network and returns the output.
+    EigenVec feed_forward(EigenVecRef input) const;
+
+    // Equivalent to feed_forward but updates the activations.
+    EigenVec feed_forward_train(EigenVecRef input);
+
+    // Trains the network using a single input-output pair.
+    // Returns the cost before training.
+    float train(EigenVecRef input, EigenVecRef output, Optimizer& optimizer);
+
+    // Getters
+    uint32_t get_layers() const;
+    std::vector<uint32_t> get_layer_sizes() const;
+    const std::vector<EigenMat>& get_weights() const;
+    const std::vector<EigenVec>& get_biases() const;
+    std::vector<float (*)(float)> get_activation_funcs() const;
+    std::vector<float (*)(float)> get_activation_derivatives() const;
+    OutputActivationFunc get_output_activation() const;
+    CostFunc get_cost() const;
+    OutputLayerDerivative get_output_layer_derivative() const;
+    const std::vector<EigenVec>& get_activations() const;
+
+  private:
+    // Helper for feed_forward functions
+    EigenVec feed_forward_helper(EigenVecRef input, bool cache) const;
+
+    // Number of layers in newtork, including input, hidden, and output layers.
+    uint32_t layers;
+    // The size of each layer.
+    std::vector<uint32_t> layer_sizes;
+    // The weights connecting each layer with the next.
+    // weights[i] shape: (layer_sizes[i] cols, layer_sizes[i+1] rows).
+    std::vector<EigenMat> weights;
+    // The biases in each hidden layer.
+    std::vector<EigenVec> biases;
+    // The activation function used in each hidden layer.
+    std::vector<float (*)(float)> activation_funcs;
+    // Derivatives of the activation functions
+    std::vector<float (*)(float)> activation_derivatives;
+    // Activation function used in output layer.
[…]

### 2025-03-23 c/src/network/network.cpp
tweaked netarch slightly

@@ -1,4 +1,3 @@
-#pragma once
 #include "network.hpp"
 #include "Eigen/Dense"
 #include "types.hpp"
@@ -49,7 +48,7 @@ Network::Network(uint32_t layers, std::vector<uint32_t> layer_sizes,
     std::uniform_real_distribution<> dis(-1.0, 1.0);
     for (int i = 0; i < layers - 1; i++) {
         EigenMat current_weights(layer_sizes[i+1], layer_sizes[i]);
-        EigenMat current_biases(layer_sizes[i+1], layer_sizes[i]);
+        EigenVec current_biases(layer_sizes[i+1], layer_sizes[i]);
         for (int col = 0; col < current_weights.cols(); col++) {
             for (int row = 0; row < current_weights.rows(); row++) {
                 current_weights(row, col) = dis(gen);
@@ -58,6 +57,8 @@ Network::Network(uint32_t layers, std::vector<uint32_t> layer_sizes,
         for (int row = 0; row < current_biases.size(); row++) {
             current_biases[row] = dis(gen);
         }
+        this->weights.push_back(current_weights);
+        this->biases.push_back(current_biases);
     }
 }
 

### 2025-03-25 c/main.cpp
changes

@@ -1,4 +1,6 @@
-#include "src/network/network.hpp"
+#include "Newton.hpp"
+#include "Optimizer.hpp"
+#include "network.hpp"
 #include <iostream>
 
 int main() { std::cout << "Hello" << std::endl; }

### 2025-03-25 c/src/network/Optimizer.hpp

@@ -1,14 +1,0 @@
-#pragma once
-#include "Eigen/Dense"
-#include "types.hpp"
-
-// Base class for optimizers
-class Optimizer {
-public:
-    // Updates the given weights and biases from their gradients.
-    virtual void update(EigenMat &weights, EigenVec &biases,
-                        const EigenMat &weight_gradients,
-                        const EigenVec &bias_gradients,
-                        const EigenVec &activations) = 0;
-    virtual ~Optimizer() = default;
-};

### 2025-03-25 c/src/network/network.cpp

@@ -1,124 +1,0 @@
-#include "network.hpp"
-#include "Eigen/Dense"
-#include "types.hpp"
-#include <vector>
-#include <stdint.h>
-#include <random>
-
-Network::Network(uint32_t layers, std::vector<uint32_t> layer_sizes,
-                 std::vector<float (*)(float)> activation_funcs,
-                 std::vector<float (*)(float)> activation_derivatives,
-                 OutputActivationFunc output_activation, CostFunc cost,
-                 OutputLayerDerivative output_layer_derivative)
-{
-    assert(
-        layers >= 2 &&
-        "network must have at least 2 layers"
-    );
-    assert(
-        layer_sizes.size() == layers &&
-        "mismatch between layers and layer_sizes"
-    );
-    for (int size : layer_sizes) {
-        assert(size > 0 && "invalid layer size of 0");
-    }
-    assert(
-        activation_funcs.size() == layers - 1 &&
-        "mismatch between layers and activations"
-    );
-    for (float (*activation)(float) : activation_funcs) {
-        assert(activation != NULL && "invalid null activation");
-    }
-    assert(output_activation != NULL && "invalid null output activation");
-    assert(cost != NULL && "invalid null cost function");
-    assert(output_layer_derivative != NULL && "invalid null output derivative");
-    this->layers = layers;
-    this->layer_sizes = layer_sizes;
-    this-> activation_funcs = activation_funcs;
-    this->activation_derivatives = activation_derivatives;
-    this->output_activation = output_activation;
-    this->cost = cost;
-    this->output_layer_derivative = output_layer_derivative;
-    this->activations.reserve(layers - 1);
-    this->weights.reserve(layers - 1);
-    this->biases.reserve(layers - 1);
-    // Randomly init weights and biases
-    std::random_device rd;
-    std::mt19937 gen(rd());
-    std::uniform_real_distribution<> dis(-1.0, 1.0);
-    for (int i = 0; i < layers - 1; i++) {
-        EigenMat current_weights(layer_sizes[i+1], layer_sizes[i]);
-        EigenVec current_biases(layer_sizes[i+1], layer_sizes[i]);
-        for (int col = 0; col < current_weights.cols(); col++) {
-            for (int row = 0; row < current_weights.rows(); row++) {
-                current_weights(row, col) = dis(gen);
-            }
-        }
-        for (int row = 0; row < current_biases.size(); row++) {
-            current_biases[row] = dis(gen);
-        }
[…]

### 2025-03-25 c/src/network/network.hpp

@@ -1,66 +1,0 @@
-#pragma once
-#include "Eigen/Dense"
-#include "types.hpp"
-#include <stdint.h>
-#include <vector>
-
-// Forward declaration of Optimizer class
-class Optimizer;
-
-class Network {
-  public:
-    // Initializes a new Network from the given parameters.
-    Network(uint32_t layers, std::vector<uint32_t> layer_sizes,
-            std::vector<float (*)(float)> activation_funcs,
-            std::vector<float (*)(float)> activation_derivatives,
-            OutputActivationFunc output_activation, CostFunc cost,
-            OutputLayerDerivative output_layer_derivative);
-
-    // Feeds the given input through the network and returns the output.
-    EigenVec feed_forward(EigenVecRef input) const;
-
-    // Equivalent to feed_forward but updates the activations.
-    EigenVec feed_forward_train(EigenVecRef input);
-
-    // Trains the network using a single input-output pair.
-    // Returns the cost before training.
-    float train(EigenVecRef input, EigenVecRef output, Optimizer& optimizer);
-
-    // Getters
-    uint32_t get_layers() const;
-    std::vector<uint32_t> get_layer_sizes() const;
-    const std::vector<EigenMat>& get_weights() const;
-    const std::vector<EigenVec>& get_biases() const;
-    std::vector<float (*)(float)> get_activation_funcs() const;
-    std::vector<float (*)(float)> get_activation_derivatives() const;
-    OutputActivationFunc get_output_activation() const;
-    CostFunc get_cost() const;
-    OutputLayerDerivative get_output_layer_derivative() const;
-    const std::vector<EigenVec>& get_activations() const;
-
-  private:
-    // Helper for feed_forward functions
-    EigenVec feed_forward_helper(EigenVecRef input, bool cache) const;
-
-    // Number of layers in newtork, including input, hidden, and output layers.
-    uint32_t layers;
-    // The size of each layer.
-    std::vector<uint32_t> layer_sizes;
-    // The weights connecting each layer with the next.
-    // weights[i] shape: (layer_sizes[i] cols, layer_sizes[i+1] rows).
-    std::vector<EigenMat> weights;
-    // The biases in each hidden layer.
-    std::vector<EigenVec> biases;
-    // The activation function used in each hidden layer.
-    std::vector<float (*)(float)> activation_funcs;
-    // Derivatives of the activation functions
-    std::vector<float (*)(float)> activation_derivatives;
-    // Activation function used in output layer.
[…]

### 2025-03-25 c/src/network/types.hpp

@@ -1,10 +1,0 @@
-#pragma once
-#include "Eigen/Dense"
-
-typedef Eigen::VectorXf EigenVec;
-typedef Eigen::Ref<EigenVec> EigenVecRef;
-typedef Eigen::MatrixXf EigenMat;
-typedef Eigen::Ref<EigenMat> EigenMatRef;
-typedef void (*OutputActivationFunc)(EigenVecRef);
-typedef float (*CostFunc)(EigenVecRef predictions, EigenVecRef labels);
-typedef EigenVec (*OutputLayerDerivative)(EigenVecRef predictions, EigenVecRef labels);

### 2025-04-03 c/src/python/network.py
This shi doesn't work

@@ -1,0 +1,45 @@
+import tensorflow as tf
+from tensorflow.keras import layers, models, utils
+from tensorflow.keras.optimizers import Optimizer
+from sklearn.datasets import load_digits
+from sklearn.model_selection import train_test_split
+from sklearn.preprocessing import StandardScaler
+import numpy as np
+
+# Load and preprocess the digits dataset
+digits = load_digits()
+X = digits.data  # shape: (n_samples, 64) since images are 8x8
+y = digits.target
+
+# Scale the features for better training
+scaler = StandardScaler()
+X_scaled = scaler.fit_transform(X)
+
+# Convert labels to one-hot encoded vectors (10 classes: digits 0-9)
+y_categorical = utils.to_categorical(y, num_classes=10)
+
+# Split the dataset into training and testing sets
+X_train, X_test, y_train, y_test = train_test_split(X_scaled, y_categorical, test_size=0.2, random_state=42)
+
+# Build the model: input layer is defined by the input shape of the first Dense layer.
+model = models.Sequential([
+    layers.Dense(128, activation='relu', input_shape=(64,)),  # Input layer + 1st hidden layer
+    layers.Dense(64, activation='relu'),                      # 2nd hidden layer
+    layers.Dense(32, activation='relu'),                      # 3rd hidden layer
+    layers.Dense(10, activation='softmax')                    # Output layer for 10 classes
+])
+
+# Instantiate your custom optimizer
+custom_optimizer = CustomOptimizer(learning_rate=0.01)
+
+# Compile the model using the custom optimizer and categorical crossentropy loss
+model.compile(optimizer=Optimizer,
+              loss='categorical_crossentropy',
+              metrics=['accuracy'])
+
+# Train the model for 100 epochs
+model.fit(X_train, y_train, epochs=100, batch_size=32, validation_split=0.1)
+
+# Evaluate the model on the test set
+test_loss, test_accuracy = model.evaluate(X_test, y_test)
+print(f"Test accuracy: {test_accuracy:.4f}")

### 2025-04-06 c/src/python/network.py
working final(?) code

@@ -1,6 +1,6 @@
 import tensorflow as tf
 from tensorflow.keras import layers, models, utils
-from tensorflow.keras.optimizers import Optimizer
+from tensorflow.keras.optimizers import Adam
 from sklearn.datasets import load_digits
 from sklearn.model_selection import train_test_split
 from sklearn.preprocessing import StandardScaler
@@ -29,11 +29,8 @@ model = models.Sequential([
     layers.Dense(10, activation='softmax')                    # Output layer for 10 classes
 ])
 
-# Instantiate your custom optimizer
-custom_optimizer = CustomOptimizer(learning_rate=0.01)
-
-# Compile the model using the custom optimizer and categorical crossentropy loss
-model.compile(optimizer=Optimizer,
+# Compile the model using an actual optimizer instance
+model.compile(optimizer=Adam(learning_rate=0.001),  # Correct optimizer usage
               loss='categorical_crossentropy',
               metrics=['accuracy'])
 

### 2025-04-06 latex/paper.tex
Added names
Small change adding names to the author tag in the paper

@@
 \documentclass{article}
-\usepackage{amsmath,amssymb}
+\usepackage[utf8]{inputenc}
+\usepackage{amsmath, amssymb}
+\usepackage{hyperref}
+\usepackage{geometry}
+\usepackage{enumitem}
 \usepackage{graphicx}
-\usepackage{enumitem}
-\usepackage{geometry}
+
 \geometry{legalpaper, portrait, margin=.5in}
+
 \title{CS 2051 Group Project}
-\author{}
-\date{March 2025}
+\author{Ibrahim Ahmad, Cipriano Dorbessan, Christopher Gossmann}
+\date{April 2025}
 
@@
 \maketitle
+
+\section*{Project Outline}
+Our project is focused on using Newton’s method (also known as the Newton-Raphson method) as the optimizer for machine learning.
+We will use \LaTeX{} to make our research paper.
+Since our project involves a significant amount of coding, we created a GitHub where we will keep our code: \url{https://github.com/cipridorbe/Newtons-Method}
+
+\section*{Background Knowledge}
+
+\subsection*{Newton’s Method}
+Newton’s method is an iterative root-finding algorithm for functions.
+For a single variable function \( f: \mathbb{R} \to \mathbb{R} \), the algorithm starts with an initial guess \( x_0 \), and the next term \( x_1 \) is found by using the tangent line at \( x_0 \).
+
+The tangent through \( (x_0, f(x_0)) \) with slope \( f'(x_0) \) is given by:
+\[
+y - f(x_0) = f'(x_0)(x - x_0)
+\]
+Solving \( y = 0 \), we get:
+\[
+x = x_0 - \frac{f(x_0)}{f'(x_0)}
+\]
+So the formula for the \(n\)-th term is:
+\[
+x_{n+1} = x_n - \frac{f(x_n)}{f'(x_n)}
+\]
+
+In machine learning, we aim to find the minimum of a function, not its root. Since the minimum occurs when \( f'(x) = 0 \), we can use Newton’s method to find the minimum:
+\[
+x_{n+1} = x_n - \frac{f'(x_n)}{f''(x_n)}
+\]
+
+For multivariate functions \( f: \mathbb{R}^k \to \mathbb{R} \), we generalize the derivative to the gradient and the second derivative to the Hessian matrix. Thus, the update formula becomes:
+\[
+x_{n+1} = x_n - [\nabla^2 f(x_n)]^{-1} \nabla f(x_n)
+\]
+Where \( x_n, x_{n+1}, \nabla f(x_n) \in \mathbb{R}^k \), and \( \nabla^2 f(x_n) \in \mathbb{R}^{k \times k} \).
+
+Computing the inverse of the Hessian matrix is expensive. Instead, we solve the linear system:
+\[
+[\nabla^2 f(x_n)] h = -\nabla f(x_n)
+\]
+Which gives \( h = -[\nabla^2 f(x_n)]^{-1} \nabla f(x_n) \).
+
+\section*{Differentiation}
+To compute derivatives, we can use:
+
+\begin{itemize}
+    \item \textbf{Finite differencing:} Simple to implement but computationally expensive when \( f''(x_n) \) is needed many times.
+    \item \textbf{Automatic differentiation:} Builds a computational graph to calculate derivatives. It is complex to implement but more efficient in higher-order systems.
+\end{itemize}
+
+\section*{Method and Complexity in Machine Learning}
+To apply Newton’s method in machine learning, we require:
+\begin{itemize}
+    \item An initial guess (often random).
+    \item A tolerance for convergence.
+    \item A maximum number of iterations.
+\end{itemize}
+
+Newton’s method converges quadratically near the solution — each iteration roughly doubles the number of correct digits.
+However, poor initial guesses may lead to slow convergence or divergence.
+The function must have continuous first and second derivatives.
+
+Performance also depends on the function shape. Linear first derivatives tend to lead to faster convergence.
+
+Although Newton’s method has a quadratic convergence rate, it can have greater complexity than other optimization algorithms due to:
+
+\begin{itemize}
+    \item Gradient computation: \( \mathcal{O}(n) \)
+    \item Hessian computation: \( \mathcal{O}(n^2) \)
+    \item Solving linear systems: \( \mathcal{O}(n^3) \) (for dense matrices)
+\end{itemize}
+
+The number of iterations varies depending on the tolerance, curvature, and initial guess.
+
+\section*{Conclusion}
+We have not yet implemented and benchmarked these algorithms, so the efficiency remains unknown.
+
+\section*{Extensions and Applications}
+Potential extensions:
+\begin{itemize}
+    \item Apply Newton’s method to other domains: physics engines, architectural design, etc.
+    \item Use in other machine learning fields: unsupervised or reinforcement learning.
+    \item Explore other optimization algorithms, including derivative-free methods.
+\end{itemize}
+
+If results are promising, we aim to test on larger networks and frameworks.
+
+\section*{References}
+\begin{enumerate}
+    \item Wikipedia, “Newton’s method.” \url{https://en.wikipedia.org/wiki/Newton%27s_method}
+    \item GeeksforGeeks, “Newton’s method in machine learning.” \url{https://www.geeksforgeeks.org/newtons-method-in-machine-learning/}
+    \item Wikipedia, “Automatic differentiation.” \url{https://en.wikipedia.org/wiki/Automatic_differentiation}
+\end{enumerate}
+
 \end{document}

### 2025-04-06 latex/paper.tex
Added the complete latex document

--- 2025-03-22 latex/paper.tex
@@
 \documentclass{article}
-\usepackage{amsmath,amssymb}
+\usepackage[utf8]{inputenc}
+\usepackage{amsmath, amssymb, amsfonts}
+\usepackage{hyperref}
+\usepackage{geometry}
+\usepackage{enumitem}
 \usepackage{graphicx}
-\usepackage{enumitem}
-\usepackage{geometry}
-\geometry{legalpaper, portrait, margin=.5in}
+\usepackage[T1]{fontenc}
+\usepackage{hyperref}
+\usepackage[version=4]{mhchem}
+\usepackage{stmaryrd}
+\usepackage{bbold}
+\usepackage[export]{adjustbox}
+\graphicspath{ {./images/} }
+
 \title{CS 2051 Group Project}
-\author{}
-\date{March 2025}
+\author{Ibrahim Ahmad, Cipriano Dorbessan, Christopher Gossmann}
+\date{April 2025}
+\hypersetup{colorlinks=true, linkcolor=blue, filecolor=magenta, urlcolor=cyan,}
+\urlstyle{same}
 
 \begin{document}
+\section*{Optimizing Newton's Method for Machine Learning}
+Our project is focused on using Newton's method (also known as the Newton-Raphson method) as the optimizer for machine learning.
+We will use LaTeX to make our research paper.
+Since our project involves a significant amount of coding, we created a Github where we will keep our code: \href{https://github.com/cipridorbe/Newtons-Method}{https://github.com/cipridorbe/Newtons-Method}
 
-\maketitle
+\section*{Background knowledge}
+\section*{Newton's method}
+Newton's method is an iterative root-finding algorithm for functions.
+For a single variable function $f: \mathbb{R} \rightarrow \mathbb{R}$ the algorithm starts with an initial guess $x_{0}$, and the next term $x_{1}$ is found by finding the root of the tangent at $x_{0}$.
+The tangent through $\left(x_{0}, f\left(x_{0}\right)\right)$ with slope $f^{\prime}\left(x_{0}\right)$ is given by $y-f\left(x_{0}\right)=f^{\prime}\left(x_{0}\right)\left(x-x_{0}\right)$, and its root occurs at $y=0$, therefore the root is $x=x_{0}-\frac{f\left(x_{0}\right)}{f^{\prime}\left(x_{0}\right)}$.
+Therefore, the formula for the nth term is
+
+$$
+x_{n+1}=x_{n}-\frac{f\left(x_{n}\right)}{f^{\prime}\left(x_{n}\right)}
+$$
+
+In machine learning, however, the goal is to find the minimum of a function, not its root.
+Since the minimum occurs when $f^{\prime}(x)=0$, we can use Newton's method to find the minimum of $f(x)$ by finding the root $f^{\prime}(x)$.
+Therefore we get the formula
+
+$$
+x_{n+1}=x_{n}-\frac{f^{\prime}\left(x_{n}\right)}{f^{\prime \prime}\left(x_{n}\right)}
+$$
+
+In machine learning, however, the function which we try to minimize is of the form $f: \mathbb{R}^{k} \rightarrow \mathbb{R}$, where $k$ is the number of parameters of the model which is (usually) large.
+To perform Newton's method in this multivariate domain, we can generalize the derivative to the gradient and the second derivative to the gradient of the gradient, which is the Hessian Matrix (this will be explained in more detail in the paper).
+Therefore we get that the formula
+
+$$
+x_{n+1}=x_{n}-\left[f^{\prime \prime}\left(x_{n}\right)\right]^{-1} f^{\prime}\left(x_{n}\right)
+$$
+
+Where $x_{n}, x_{n+1}$ and $f^{\prime}\left(x_{n}\right)$ are vectors in $\mathbb{R}^{k}$ and $f^{\prime \prime}\left(x_{n}\right)$ is a real matrix of size $k \times k$.
+Finding the inverse of such a large matrix can often be an expensive operation.
+Since $-\left[f^{\prime \prime}\left(x_{n}\right)\right]^{-1} f^{\prime}\left(x_{n}\right)$ is a vector in $\mathbb{R}^{k}$, we can find this vector by solving the linear system:
+
+$$
+h=-\left[f^{\prime \prime}\left(x_{n}\right)\right]^{-1} f^{\prime}\left(x_{n}\right) \Rightarrow\left[f^{\prime \prime}\left(x_{n}\right)\right] h=-f^{\prime}\left(x_{n}\right)
+$$
+
+\begin{center}
+\includegraphics[max width=\textwidth]{2025_04_06_546e56f61867917283deg-2}
+\end{center}
+
+Newton's method visualization for $f(x)=x^{4}+.1$
+
+\section*{Differentiation}
+To find the derivative $f^{\prime \prime}\left(x_{n}\right)$ numerous algorithms can be used. The easiest to implement is finite differencing where
+
+$$
+f^{\prime \prime}(x) \approx \frac{\frac{f(x+h)-f(x)}{h}-\frac{f(x)-f(x-h)}{h}}{h}
+$$
+
+However, this becomes computationally expensive when $f^{\prime \prime}\left(x_{n}\right)$ is used numerous times across multiple $x_{n}$.
+A more difficult method to implement is Automatic Differentiation, where a tree is built containing the function and its derivatives, and is traversed multiple times at runtime.
+Computation of the derivative this way is more complex but is faster in higher order systems, as fewer matrix operations are required.
+
+\section*{Method and complexity in Machine Learning}
+In applications for machine learning, Newton's method optimization requires the following data: an initial guess for the minimum of the function, a tolerance level for convergence, and a maximum number of iterations.
+Here, the initial guess is randomly generated, but the further the guess is from the actual minimum, the longer it may take for Newton's method to find the minimum.
+
+Newton's method converges to a solution at least quadratically (this holds only for the neighborhood of the root), in that each iteration the number of accurate digits approximately doubles.
+Nevertheless, this can be affected by a variety of factors.
+As mentioned in the previous paragraph, if the initial guess is very close to the solution, the number of iterations may be very few, but if it's very far, it may not only cause slow convergence, but even divergence and never find the minimum of the function.
+Additionally, Newton's method, because it's based on differentiation, relies on the function to have continuous first and second derivatives, so it is very inefficient or cannot be used at all if the function has discontinuities or singularities.
+Lastly, Newton's method ends when the change in the predicted solution is small enough or when the value of the prediction is very close to 0 .
+Newton's method's performance may also vary depending on the shape of the function.
+Functions with a relatively linear first derivative may have faster convergence in the input space, as the tangent line approximation of the first derivative by the second derivative is more accurate.
+
+Because Newton's method has a quadratic convergence rate, it is considered to be a good choice based on the number of iterations needed to find the solution, however it can still have greater complexity than other optimization algorithms.
+This is mainly because each iteration requires calculating both the gradient (first derivative) and the Hessian (second derivative), so it can be very computationally expensive when used with a lot of variables.
+Also, the manipulation and storage of the Hessian matrix is very memory-expensive, especially for high dimensions.
+With $n$ variables, calculating the gradient is $O(n)$ and calculating the Hessian is $O\left(n^{2}\right)$.
+Lastly, solving the linear systems with any method is usually $O\left(n^{3}\right)$ when it's dense.
+The number of iterations can vary greatly, as it depends on the tolerance level, curvature of the function, and the initial guess.
+
+\section*{Conclusion}
+As we have not implemented these algorithms and benchmarked them against various types of problems, we do not know the efficiency of these algorithms yet.
+
+\section*{Extensions and Applications}
+Extensions include using Newton's method as the optimizer for other problems, such as physics engines, architecture design, or even other machine learning related applications not covered in our project (such as unsupervised or reinforcement learning).
+Other extensions include the use of other optimization algorithms for machine learning applications, perhaps most interestingly derivative-free algorithms (non-gradient based).
+For future work, if our findings suggest Newton's method can be useful, we could continue testing using larger networks and other frameworks.
+The applications of our work depend on the findings, but will all be related to machine learning.
+
+\section*{References}
+[1] "Newton's method," Wikipedia, \href{https://en.wikipedia.org/wiki/Newton%27s_method}{https://en.wikipedia.org/wiki/Newton's\_method} (accessed Mar. 7, 2025).\\[0pt]
+[2] GeeksforGeeks, "Newton's method in machine learning," GeeksforGeeks, \href{https://www.geeksforgeeks.org/newtons-method-in-machine-learning/}{https://www.geeksforgeeks.org/newtons-method-in-machine-learning/} (accessed Mar.
+7, 2025).\\[0pt]
+[3] "Automatic differentiation," Wikipedia, \href{https://en.wikipedia.org/wiki/Automatic_differentiation}{https://en.wikipedia.org/wiki/Automatic\_differentiation} (accessed Mar. 7, 2025).
+
 \end{document}

### 2025-04-07 c/src/python/hessian.py
working

@@ -1,12 +1,15 @@
+import os
+import csv
+import random
+import numpy as np
+import matplotlib.pyplot as plt
 import tensorflow as tf
 from tensorflow.keras import layers, models, utils
 from sklearn.datasets import load_digits
 from sklearn.model_selection import train_test_split
 from sklearn.preprocessing import StandardScaler
-import numpy as np
-import matplotlib.pyplot as plt
 
-EPOCHS = 20
+EPOCHS = 50
 
 # Load and preprocess data
 digits = load_digits()
@@ -82,68 +85,139 @@ def train_newton(model, X_train, y_train, epochs=EPOCHS, batch_size=32):
         print(f"Newton Epoch {epoch+1}: Loss={history['loss'][-1]:.4f}, Accuracy={history['acc'][-1]:.4f}")
     return history
 
-# Training with Adam
+# Adam
 def train_adam(X_train, y_train, X_test, y_test):
     model = build_model()
     model.compile(optimizer=tf.keras.optimizers.Adam(learning_rate=0.001),
-                  loss='categorical_crossentropy',
-                  metrics=['accuracy'])
+                  loss='categorical_crossentropy', metrics=['accuracy'])
     history = model.fit(X_train, y_train, validation_data=(X_test, y_test),
                         epochs=EPOCHS, batch_size=32, verbose=1)
     return history, model
 
-# Training with plain SGD (gradient descent)
+# SGD
 def train_sgd(X_train, y_train, X_test, y_test):
     model = build_model()
     model.compile(optimizer=tf.keras.optimizers.SGD(learning_rate=0.01),
-                  loss='categorical_crossentropy',
-                  metrics=['accuracy'])
+                  loss='categorical_crossentropy', metrics=['accuracy'])
     history = model.fit(X_train, y_train, validation_data=(X_test, y_test),
                         epochs=EPOCHS, batch_size=32, verbose=1)
     return history, model
 
+# Utilities for model saving/loading and CSV
+
+def save_model(model, name):
+    model.save(f"{name}_model.h5")
+
+def load_model(name):
+    return tf.keras.models.load_model(f"{name}_model.h5")
+
+def write_history_to_csv(histories, filename="results.csv"):
[…]

### 2025-04-08 latex/Draft/Draft.tex
Added draft latex

\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage[export]{adjustbox}
\graphicspath{ {./images/} }
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[version=4]{mhchem}
\usepackage{stmaryrd}

\title{Optimizing Newton's Method for Machine Learning }

\author{}
\date{}

\begin{document}
\maketitle
\section*{Introduction}
In this project, we explored and tested the practical effectiveness of Newton's Method as an optimization technique within the machine learning context. We were interested in investigating whether this second-order method could be effectively adapted and optimized to train neural networks on a simple supervised learning problem: digit classification using the digits dataset.

To do this, we developed a feedforward neural network and trained it on the digits dataset provided by scikit-learn. What we focused our project on is the implementation of our own Newton's Method optimizer, including the computation of gradients and Hessians, and solving the resulting linear systems at each iteration. For this we had to learn the mathematical foundations of Newton's Method and numerical differentiation, matrix manipulation, and computational efficiency.

Once we had a working implementation, we compared the performance of our Newton-based optimizer against two widely used optimization algorithms in machine learning: Stochastic Gradient Descent (SGD) and ADAM. By training and testing the model with the different optimizers, we observed that our optimizer was able to reduce training error over time and showed the convergence properties we were looking for. However, despite these results we achieved, our optimizer did not reach the level of performance achieved by ADAM or SGD in terms of accuracy, and also our optimizer took a lot more time to train when compared to these others.

This outcome shows that using Newton's Method for machine learning is possible, but might not be ideal. While it also uses the second-order derivative to converge, it is made worse because of the computational complexity of calculating the Hessian matrix-especially in higher dimensions. However, this project can serve as insight on the potential challenges of incorporating second-order optimization techniques into neural network training.

\section*{Newton's method in optimization}
Missing text goes here

\section*{Network model}
We used Tensorflow with Keras to build our network model. The dataset we used was "Pen-Based Recognition of Handwritten Digits" which consists of 8x8 images of handwritten digits. Our network consists of an input layer of 64 neurons (one for each pixel), a hidden layer of 32 neurons with ReLU as their activation function, and an output layer of 10 neurons (one for each digit) with softmax as the activation. For the ADAM and SGD optimizers, we used Keras' built-in optimizers. We wrote the Newton's Method optimizer ourselves using Keras' automatic differentiation (through the use of Gradient Tapes and the built-in Jacobian method). The Adam and SGD models were trained using Tensorflow's fit method, and the Newton's method model was trained using our own train function, as there was no pre-existing function to do this. When analyzing the results, it is important to take into account that the networks using ADAM and SGD will have a slight advantage over the network using our custom Newton's Method optimizer, as the code for these two optimizers was written and optimized by Tensorflow developers over many years.

\section*{Results and conclusions}
We trained and evaluated the model's performance over the course of 50 training epochs, comparing our Newton's Method optimizer with ADAM and Stochastic Gradient Descent (SGD). The results, shown in the training loss and accuracy plots, reveal several important insights into the strengths and limitations of each approach.\\
\includegraphics[max width=\textwidth, center]{2025_04_08_c5129b05008e68f8b3cdg-2}

From the training loss graph, it can be seen that ADAM achieved the fastest and most stable convergence, reducing the loss to near zero within the first 20 epochs. SGD also performed well, it still converged slower than ADAM but still significantly reduced the loss over the epochs. Newton's Method, while improving the loss steadily, underperformed when compared to both ADAM and SGD, as it showed a slower rate of convergence.

Similarly, the training accuracy graph highlights that ADAM quickly pushed the model to over $95 \%$ accuracy, with SGD really close to it. Newton's Method displayed a more gradual improvement in accuracy, reaching around $93 \%$ by the end of training. While this is a very good result, it always performed worse than the other two optimizers.

These results suggest that while Newton's Method has theoretical advantages, such as quadratic convergence near the optimal values, its practical application in training neural networks is limited by the computational cost of its operations and because the initial conditions affect it greatly. The cost of calculating the Hessian matrix at each step likely contributed to its slower performance.

Despite this, our implementation demonstrated that Newton's Method is capable of reducing error and improving accuracy just by calculating the gradient and the hessian, and without using any adaptive strategies like ADAM.

In conclusion, our project shows that Newton's Method works as an optimization algorithm in machine learning but may not be ideal, especially when compared with modern first-order optimizers for large-scale and real-time training tasks, as these take much less time than our custom Newton's method optimizer.

\section*{Extensions and Applications}
Once again, the performance of the Newton's Method optimizer was heavily impacted by the fact that we wrote the optimizer ourselves, rather than using a pre-written optimizer. A direct extension on our project would be to optimize our optimizer. One such way to do this is to code it in a faster language like $\mathrm{C} / \mathrm{C}++$ and to improve on the calculation of the hessian.

Other extensions include using Newton's Method in other machine learning contexts, such as non-classification neural networks such as regression, or even in unsupervised or reinforcement learning. Another possible extension is the use of other optimization algorithms to train networks, such as derivative-free algorithms. We could also use Newton's Method in other contexts, such as in symbolic calculators.

\section*{References}

\end{document}

### 2025-04-09 latex/Draft/Draft.tex
saving progress

@@
 
-\title{Optimizing Newton's Method for Machine Learning }
+\title{Newton's Method in Machine Learning }
 
-\author{}
-\date{}
+\author{Ibrahim Ahmad, Cipriano Dorbessan, Christopher Gossmann}
+\date{April 2025}
 
@@
 
-\section*{Newton's method in optimization}
-Missing text goes here
+\section*{Proofs}
+There are a few nontrivial statements that are crucial for understanding Numerical Optimization and training Neural Networks.
+\subsection*{Proof Of Taylor's Theorem}
+Taylor's Theorem is cruicial for doing numerical optimization. It allows us to provide an approximation of a function in the open neighbourhood of a point.
+
+We must prove Taylor's theorem for the high dimensional case. We shall do so directly.
+Let $f : \mathbb{R}^n \mapsto \mathbb{R}$.
+We shall observe some properties of $f$ in the region of $A$, where $A$ is an open subset of $\mathbb{R}^n$.
+We shall make the following assumptions: $f$ is continuous on $A$, $f$ is differentiable on $A$.
+(Note: We do not contain the boundary points as this allows us to handle asymptotics more neatly).
+Let us define the derivative of $f$ in the direction of $\alpha$ by writing \begin{equation}\vec{\alpha} \cdot \vec{\nabla} f(\vec{x}) = \lim_{t \to 0}\frac{f(\vec{x} + \vec{\alpha}t) - f(\vec{x})}{t}\end{equation}, under the assumption that $\vec{x} + t \vec{\alpha} \in A$.
+This allows us to find the derivative in any direction from the location of $x$.
+Let us define $\Phi(t) = f(\vec{x} + \vec{\alpha} t)$.
+Because $\Phi : \mathbb{R} \mapsto \mathbb{R}$
+we can use Taylor's theorem on $\Phi$. This lets us write the equivalency 
+\begin{equation}\Phi(t) = \Phi(0) + \Phi'(0)t + \Phi''(0)\frac{t^2}{2} + \dots\end{equation}
+By substituting the definition of $\Phi(t)$ into both sides, we can write 
+\begin{equation} f(\vec{x} + t \cdot \vec{\alpha}) = \Phi(0) + \Phi'(0)t + \Phi''(0)\frac{t^2}{2} + \dots\end{equation}
+We also know that 
+\begin{equation}\Phi(t)' = \lim_{t \to 0}\frac{f(\vec{x} + \vec{\alpha}t) - f(\vec{x})}{t} = \vec{\alpha} \cdot \vec{\nabla}f(\vec{x})\end{equation}
+By the definition of $\Phi$, and $\vec{\alpha} \cdot \vec{\nabla}$.
+By substituting into both sides, we get the following. For convenience, we shorten $\vec{\alpha}\cdot\vec{\nabla}$ to $\vec{\nabla}$.
+\begin{equation} f(\vec{x} + t \cdot \vec{\alpha}) = f(\vec{x}) + [\vec{\nabla}f(\vec{x})]^T\vec{\alpha} t + \frac{1}{2}\vec{\alpha}^T[\vec{\nabla}^2f(\vec{x})]\vec{\alpha} \cdot t^2 + \dots\end{equation}
+We shall only use first and second order methods, so we have not shown the third order term of the Taylor Series, which requires rank 3 tensor multiplications.
+Thus we have proved Taylor's Theorem for functions $f : \mathbb{R}^n \mapsto \mathbb{R}$, under an open subset of $\mathbb{R}^n$.
+
+\subsection*{Convergence Of Gradient Descent}
+While we do not use Gradient Descent to train any of the models, understanding Gradient Descent and it's properties is crucial to understanding Adam, Stochastic Gradient Descent, and Newton's Method.
+For further reading regarding these algorithms we reccomend \cite{garrigos_2023_handbook}.
+
+Because the Neural Network is not L-Smooth, or Convex, we must show the convergence of Gradient Descent under minimal assumptions.
+Let $x_k$ be the value of the descent's algorithm's value at the $k$'th iteration.
+We shall use the following update rule $x_{k+1} = x_k - \eta_k\vec{\nabla}f(x_k)$.
+
+By the mean value theorem, we know that $f(\vec{x_{k+1}}) = f(\vec{x_k}) - \eta \|\vec{\nabla}f(x_k)\|^2 + o(\eta\|\vec{\nabla}f(\vec{x_k}\|^2)$.
+The proof of the $o$ term is complex and not useful for what we intend to do.
+It can be imagined as an upper bound of the remainder of the Taylor approximation $f(x_{k+1})$ after the first term.
+If $\eta_k$ is sufficently small, we can write $f(x_{k+1}) \leq f(x_k) - \eta_k \|\vec{\nabla}f(x_k)\|^2$.
+We must make the assumption that $\eta_k$ is sufficently small, as without that assumption we may overshoot the point we intend on converging to.
+We note, that under these assumptions $f(x_{k+1}) < f(x_k)$ if $\eta_k \|\vec{\nabla}f(x_k)\|^2$ is nonzero, and $f(x_{k+1}) = f(x_k)$ if $\eta_k \|\vec{\nabla}f(x_k)\|^2$ is zero.
+Thus we have shown that the function values are decreasing, and the algorithm converges at a minima or saddle point if $f$ is lower bounded.
+
+We note that $\lim_{k \to \infty} \|\vec{\nabla}\| = 0$, as that is the state the algorithm must converge to to be considered to have converged.
+We also know that $\eta_k > 0$, as if $\eta_k = 0$, then the algorithm would not converge.
+
+\subsection*{Convergence Of Other Descent Algorithms}
+
+We shall now study the convergence of algorithms of the following form $x_{k+1} = x_k - S_k\vec{\nabla}f(x_k)$.
+The constraints we place are that $S_k\vec{\nabla}f(x_k)$ is sufficiently small as to not overshoot the extrema, and that $S_k$ is a positive definite matrix.
+For example, Newton's method chooses $S_k = \vec{\nabla}^{-2}f(x_k)$.
+
+Algorithms that follow this form are guarunteed to converge to a minima such that $\vec{\nabla}f(x_k) = 0$.
+Recall the following property of positive definite matrices, for some arbitrary vector $\vec{v}$, $\vec{v}^TS\vec{v} > 0$.
+In other words, when multiplying the matrix $S$ by some vector, that vector is rotated by less than $90^{\circ}$.
+Thus we know that $(S_k\vec{\nabla}f(x_k)) \cdot \vec{\nabla}f(x_k) > 0$.
+Let us now project $S_k\vec{\nabla}f(x_k)$ onto $\vec{\nabla}f(x_k)$, yielding $d_k$.
+Because $S_k$ is semi definite, we know that $d_k$ is nonzero and $d_k \cdot ,\vec{\nabla}f(x_k)$ or in other words, $S_k$ transforms the gradient vector such that it no longer faces the direction of steepest descent, but it still faces a direction of descent.
+Because the projection in the descent direction is a nonzero positive scaled version of the gradient, the descent can be considered to be similar to that of gradient descent with a different stepsize.
 
@@
 
-\section*{References}
+\bibliographystyle{IEEEtran}
+\bibliography{main}
 

### 2025-04-09 latex/Draft/Draft.tex
new paper

@@
 \maketitle
-\section*{Introduction}
+\section{Introduction}
 In this project, we explored and tested the practical effectiveness of Newton's Method as an optimization technique within the machine learning context.
@@
 
-Once we had a working implementation, we compared the performance of our Newton-based optimizer against two widely used optimization algorithms in machine learning: Stochastic Gradient Descent (SGD) and ADAM.
+Once we had a working implementation, we compared the performance of our Newton-based optimizer against two widely used optimization algorithms in machine learning: Stochastic Gradient Descent (SGD)\cite{contributors-2020-stochastic} and ADAM \cite{kingma2017adammethodstochasticoptimization}.
 By training and testing the model with the different optimizers, we observed that our optimizer was able to reduce training error over time and showed the convergence properties we were looking for.
@@
 However, this project can serve as insight on the potential challenges of incorporating second-order optimization techniques into neural network training.
+\newpage
+\subsection{Background of Newton's Method and complexity in Machine Learning}
+Previously we knew that Newton’s method optimization requires the following data: an initial guess for the minimum of the function, a tolerance level for convergence, and a maximum number of iterations.
+Here, the initial guess is randomly generated, but the further the guess is from the actual minimum, the longer it may take for Newton’s method to find the minimum.
 
-\section*{Proofs}
+Newton’s method converges to a solution at least quadratically (this holds only for the neighborhood of the root)\cite{geeksforgeeks-2024-newtons}, in that each iteration the number of accurate digits approximately doubles.
+Nevertheless, this can be affected by a variety of factors.
+As mentioned in the previous paragraph, if the initial guess is very close to the solution, the number of iterations may be very few, but if it’s very far, it may not only cause slow convergence, but even divergence and never find the minimum of the function.
+Additionally, Newton’s method, because it’s based on differentiation, relies on the function to have continuous first and second derivatives, so it is very inefficient or cannot be used at all if the function has discontinuities or singularities.
+Lastly, Newton’s method ends when the change in the predicted solution is small enough or when the value of the prediction is very close to 0.
+Newton’s method’s performance may also vary depending on the shape of the function.
+Functions with a relatively linear first derivative may have faster convergence in the input space, as the tangent line approximation of the first derivative by the second derivative is more accurate.
+
+Because Newton’s method has a quadratic convergence rate, it is considered to be a good choice based on the number of iterations needed to find the solution, however it can still have greater complexity than other optimization algorithms.
+This is mainly because each iteration requires calculating both the gradient (first derivative) and the Hessian (second derivative), so it can be very computationally expensive when used with a lot of variables.
+Also, the manipulation and storage of the Hessian matrix is very memory-expensive, especially for high dimensions.
+With n variables, calculating the gradient is $O(n)$ and calculating the Hessian is $O(n^2)$.
+Lastly, solving the linear systems with any method is usually $O(n^3)$ when it’s dense.
+The number of iterations can vary greatly, as it depends on the tolerance level, curvature of the function, and the initial guess.
+
+\section{Proofs}
 There are a few nontrivial statements that are crucial for understanding Numerical Optimization and training Neural Networks.
-\subsection*{Proof Of Taylor's Theorem}
-Taylor's Theorem is cruicial for doing numerical optimization. It allows us to provide an approximation of a function in the open neighbourhood of a point.
+\subsection{Proof Of Taylor's Theorem}
+Taylor's Theorem is cruicial for doing numerical optimization \cite{recht-wright}. It allows us to provide an approximation of a function in the open neighbourhood of a point.
 
@@
 Because $\Phi : \mathbb{R} \mapsto \mathbb{R}$
-we can use Taylor's theorem on $\Phi$. This lets us write the equivalency 
+we can use the one dimensional form of Taylor's theorem \cite{wikipediacontributors-2019-taylor}on $\Phi$. This lets us write the equivalency 
 \begin{equation}\Phi(t) = \Phi(0) + \Phi'(0)t + \Phi''(0)\frac{t^2}{2} + \dots\end{equation}
@@
 
-\subsection*{Convergence Of Gradient Descent}
+\subsection{Convergence Of Gradient Descent}
 While we do not use Gradient Descent to train any of the models, understanding Gradient Descent and it's properties is crucial to understanding Adam, Stochastic Gradient Descent, and Newton's Method.
-For further reading regarding these algorithms we reccomend \cite{garrigos_2023_handbook}.
+For further reading regarding these algorithms we reccomend \cite{recht-wright}  and \cite{wikipediacontributors-2019-newtons}.
 
-Because the Neural Network is not L-Smooth, or Convex, we must show the convergence of Gradient Descent under minimal assumptions.
+Because the Neural Network is not L-Smooth\cite{lipschitz}, or Convex, we must show the convergence of Gradient Descent under minimal assumptions.
 Let $x_k$ be the value of the descent's algorithm's value at the $k$'th iteration.
@@
 
-By the mean value theorem, we know that $f(\vec{x_{k+1}}) = f(\vec{x_k}) - \eta \|\vec{\nabla}f(x_k)\|^2 + o(\eta\|\vec{\nabla}f(\vec{x_k}\|^2)$.
-The proof of the $o$ term is complex and not useful for what we intend to do.
+By the mean value theorem \cite{wikipediacontributors-2019-mean}, we know that $f(\vec{x_{k+1}}) = f(\vec{x_k}) - \eta \|\vec{\nabla}f(x_k)\|^2 + o(\eta\|\vec{\nabla}f(\vec{x_k}\|^2)$.
+The proof of the $o$ term is complex and not useful for what we intend to do, but is discussed in .
 It can be imagined as an upper bound of the remainder of the Taylor approximation $f(x_{k+1})$ after the first term.
@@
 
-\subsection*{Convergence Of Other Descent Algorithms}
+\subsection{Convergence Of Other Descent Algorithms}
 
@@
 
-\section*{Network model}
+\section{Network model}
 We used Tensorflow with Keras to build our network model.
@@
 
-\section*{Results and conclusions}
+\section{Results and conclusions}
 We trained and evaluated the model's performance over the course of 50 training epochs, comparing our Newton's Method optimizer with ADAM and Stochastic Gradient Descent (SGD).
@@
 
-\section*{Extensions and Applications}
+\section{Extensions and Applications}
 Once again, the performance of the Newton's Method optimizer was heavily impacted by the fact that we wrote the optimizer ourselves, rather than using a pre-written optimizer.
@@
 Another possible extension is the use of other optimization algorithms to train networks, such as derivative-free algorithms.
-We could also use Newton's Method in other contexts, such as in symbolic calculators.
+Newton's Method is also used for Interior-Point solvers\cite{ipopt} like Sleipnir \cite{sleipnir}.
 

### 2025-04-09 latex/Draft/Draft.tex
Changes
changes

@@
 In this project, we explored and tested the practical effectiveness of Newton's Method as an optimization technique within the machine learning context.
-We were interested in investigating whether this second-order method could be effectively adapted and optimized to train neural networks on a simple supervised learning problem: digit classification using the digits dataset.
+We were interested in investigating whether this second-order method could be effectively adapted and optimized to train neural networks on a simple supervised learning problem: digit classification using the digits dataset \cite{digits}.
 
@@
 We also know that 
-\begin{equation}\Phi(t)' = \lim_{t \to 0}\frac{f(\vec{x} + \vec{\alpha}t) - f(\vec{x})}{t} = \vec{\alpha} \cdot \vec{\nabla}f(\vec{x})\end{equation}
+\begin{equation}\Phi'(t) = \lim_{t \to 0}\frac{f(\vec{x} + \vec{\alpha}t) - f(\vec{x})}{t} = \vec{\alpha} \cdot \vec{\nabla}f(\vec{x})\end{equation}
 By the definition of $\Phi$, and $\vec{\alpha} \cdot \vec{\nabla}$.
@@
 \begin{equation} f(\vec{x} + t \cdot \vec{\alpha}) = f(\vec{x}) + [\vec{\nabla}f(\vec{x})]^T\vec{\alpha} t + \frac{1}{2}\vec{\alpha}^T[\vec{\nabla}^2f(\vec{x})]\vec{\alpha} \cdot t^2 + \dots\end{equation}
+Note: The Hessian $\vec{\nabla}^2f(\vec{x})$ appears here because it is the multidimensional equivalent of $\Phi''$.
+In other words, the Hessian appears here because it encodes the second order information required by the Taylor Polynomial.
 We shall only use first and second order methods, so we have not shown the third order term of the Taylor Series, which requires rank 3 tensor multiplications.
@@
 The constraints we place are that $S_k\vec{\nabla}f(x_k)$ is sufficiently small as to not overshoot the extrema, and that $S_k$ is a positive definite matrix.
-For example, Newton's method chooses $S_k = \vec{\nabla}^{-2}f(x_k)$.
+For example, Newton's method chooses $S_k = \vec{\nabla}^{-2}f(x_k)$ \cite{recht-wright}.
 
@@
 We used Tensorflow with Keras to build our network model.
-The dataset we used was "Pen-Based Recognition of Handwritten Digits" which consists of 8x8 images of handwritten digits.
+The dataset we used was "Pen-Based Recognition of Handwritten Digits"\cite{digits} which consists of 8x8 images of handwritten digits.
 Our network consists of an input layer of 64 neurons (one for each pixel), a hidden layer of 32 neurons with ReLU as their activation function, and an output layer of 10 neurons (one for each digit) with softmax as the activation.
